---
title: "DRY coding"
author: "Data Science in Biomedicine - Week 5"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```
## Introduction

We'll be using an in-built data set from the MASS package. But, because the MASS package can be a bit problematic to install, I've provided the data for you. It's in .rds format, so we load it in using the `read_rds()` function.

For more info about the dataset, head to https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/survey.html

```{r}
library(tidyverse)
library(knitr)
install.packages("kableExtra")
library(kableExtra)

survey <- read_rds("./data/survey.rds")
```

## Exercise 1
Explore the data - what variables are in there, what variable types, how many variables and observations?
```{r}
str(survey)
head(survey)

# 237 obs. of  12 variables
```


## Exercise 2
Round every numeric variable to 0 decimal places.

```{r}
rounded_data <- survey %>% 
  mutate(across(where(is.numeric), round)) 
  

view(rounded_data)

# round function automatically rounds to 0
```


## Exercise 3
Using the rounded dataframe you created in Exercise 2, calculate the mean of each numeric variable in the dataset. Use kable() to insert a nicely formatted table into your R Markdown document.

```{r}
mean_data <- rounded_data %>% 
  summarise(across(where(is.numeric), mean, na.rm = TRUE))

nice_table <- mean_data %>% 
  kable(
    col.names = c("Dominant Hand Span", "Non-Dominant Hand Span", "Pulse", "Height", "Age"),
    digits = 0,
    caption = "Mean measurements of student characteristics at the University of Adelaide",
    align = "ccccc") %>% 
kable_styling("striped",
              full_width = F,
              position = "center",
              font_size = "7",
              stripe_color = "gray!10")

# footnote("Data ")
                



nice_table

```


## Exercise 4
Now make exactly the same table, using the pivot functions. Again, use kable() to insert a nicely formatted table into your R Markdown document.
```{r}
pivot_table_nice <- rounded_data %>% 
  select(where(is.numeric)) %>% 
  pivot_longer(cols = everything()) %>% 
  group_by(name) %>% 
  summarise(mean = mean(value, na.rm = TRUE)) %>% 
  pivot_wider(names_from = name, values_from = mean) %>% 
  kable(col.names = c("Age", "Height", "ND Hand", "Pulse", "D Hand"),
    digits = 0,
    caption = "Mean measurements of student characteristics at the University of Adelaide",
    align = "ccccc")

pivot_table_nice




```

## Exercise 5
Students had initially expressed their height in either imperial or metric units, but someone converted everything into centimeters. Create a new variable, Height_original, which will show height in centimeters for students who expressed their height in metric units, and height in inches for students who prefer imperial units. 

NOTE: This one is a bit of a challenge. You may want to check out the `case_when()` function.

```{r}
og_units <- rounded_data %>% 
  select(Height, M.I) %>% 
  mutate(Height_original = case_when(
    M.I == "Imperial" ~ Height*0.394,
    M.I == "Metric" ~ Height
   )) %>% 
  drop_na()

view(og_units)



```


## Exercise 6
Create a table that shows the mean and standard deviation of height - in centimeters for students who used metric units, and in inches for students who used imperial units. Use kable() to insert a nicely formatted table into your R Markdown document.

```{r}
mean_and_sd <- og_units %>% 
  mutate(mean = mean(Height_original))


mean_and_sd

```


## Exercise 7
Create a contingency table to see the relationship between exercise and smoking. Pivot your table to make it easier to read, and then use kable() to insert a nicely formatted table into your R Markdown document.

```{r}

```


## Exercise 8
Create a grid of Pulse histograms - one histogram for each combination of Sex and Exercise.

```{r}

```


## Exercise 9
Create a new variable, Hnd.diff, that shows the difference in span between the writing and non-writing hand. Create a set of boxplots of this variable, with a separate boxplot for those who write with their right hand, and another one for those who write with their left hand.

```{r}

```

